Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-10 paper.tex
finished a first draft of Sectron 4.2 Refining the specification

\subsection{Infrastructure state transition}
\label{sec:statetrans}
The state transition relation uses the syntactic infix notation 
\texttt{I \ttrelI\, I'}  to denote that infrastructures 
\texttt{I} and \texttt{I'} are in this relation.
%Rules for the put, get, move, process and delete actions exist.
To give an impression of this definition, we show first the rule defining the state transition for
the action get. Initially, this rule expresses that
an actor that resides at a location \texttt{l} (\texttt{a \ttatI\ l})
and is enabled by the local policy in this location to ``get'' can
combine  all ids at the current location (contained in \texttt{egra G l}) with
all actors at the current location (contained in \texttt{agra G l}) and add this
set of pairs to his knowledge set \texttt{kgra G} using the function update \texttt{f(l := n)}
redefining the function \texttt{f} for the input \texttt{l} to have now the new value \texttt{n}.
\begin{ttbox}
 {\bf{get\_data}}: G = graphI I \ttImp a \ttatI l \ttImp enables I l (Actor a) get \ttImp 
   I' = Infrastructure 
          (Lgraph (gra G)(agra G)(cgra G)(lgra G)(egra G)
            ((kgra G)((Actor a) := ((kgra G (Actor a))(l:=
                 \{(x,y). x \ttin agra G l \ttand y \ttin egra G l\})))))
          (delta I)
   \ttImp I \ttrel{n} I' 
\end{ttbox}
Another interesting rule for the state transition is the one for \texttt{move} whose structure
resembles the previous one.
\begin{ttbox}
 {\bf{move}}: G = graphI I \ttImp  a \ttin actors_graph(graphI I) \ttImp
    l \ttin nodes G \ttImp l' \ttin nodes G \ttImp enables I l' (Actor h) move \ttImp
    I' = Infrastructure (move_graph_a a l l' (graphI I))(delta I)
 \ttImp I \ttrelI I' 
\end{ttbox}
The semantics of this rule is embedded in the function \texttt{move\_graph\_a}
that adapts the infrastructure state so that the moving actor \texttt{a} is now
associated to the target location \texttt{l'} in the actor map \texttt{agra} and not
any more at \texttt{l} and also the association of \texttt{efids} is updated accordingly.
\begin{ttbox}
move_graph_a n l l' g \ttequiv
  Lgraph (gra g) 
         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then 
           ((agra g)(l := (agra g l) - \{n\}))(l' := (insert n (agra g l')))
          else (agra g))
         (cgra g)(lgra g)
         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then
               ((egra g)(l := (egra g l) - \{efemid (cgra g n)\}))
                        (l' := (insert (efemid (cgra g n))(egra g l')))
          else egra g)
         (kgra g)
\end{ttbox}  
Based on this state transition and the above defined \texttt{Corona\_scenario}, we define the first Kripke structure.
\begin{ttbox}
 corona_Kripke \ttequiv Kripke \{ I. Corona_scenario \ttrelIstar I \} \{Corona_scenario\}
\end{ttbox}

@@

Since we consider a predicate transformer semantics, we use
sets of states to represent properties. 
%For example, the attack property
The invalidated global policy is given by the following set \texttt{shc}.
\begin{ttbox}
 scorona \ttequiv \{x. \ttexists n. \ttneg global_policy x (Efid n)\}
\end{ttbox}
The attack we are interested in is to see whether for the scenario
\begin{ttbox}
 Kripke\_scenario \ttequiv  Infrastructure ex_graph local_policies 
\end{ttbox}
from the initial state \texttt{Icorona \ttequiv \{corona\_scenario\}},
the critical state \texttt{scorona} can be reached,
that is, is there a valid attack \texttt{(Icorona,scorona)}?

For the Kripke structure \texttt{corona\_Kripke}
we first derive a valid and-attack using the attack tree proof calculus.
\begin{ttbox}
  \ttvdash [\ttcalN{(Icorona,Corona)},\ttcalN{(Corona,Corona1)},\ttcalN{(Corona1,Corona2)},
             \ttcalN{(Corona2,Corona3)},\ttcalN{(Corona3,scorona)}]\ttattand{\texttt{(Icorona,scorona)}}
\end{ttbox}
The sets \texttt{Corona, Corona1, Corona2, Corona3} are the intermediate states where
\texttt{Bob} moves to shop and \texttt{Eve} follows him collecting the Ephemeral Ids in each location.
The collected information enables to identify Bob's Ephemeral Id.

@@

\subsection{Refining the Specification}
\label{sec:corref}
Clearly, fixed Ephemeral Ids are not really ephemeral. The model presented
in Section \ref{sec:model} is deliberately been chosen very abstractly to allow focusing on
the basic system architecture and finding an initial deanonymization attack.
We now introduce ``proper'' Ephemeral Ids and show how the system datatype can be refined to a system
that uses those instead of the fixed ones.

@@

>>>>

\subsection{Infrastructure state transition}
\label{sec:statetrans}
The state transition relation uses the syntactic infix notation 
\texttt{I \ttrelI\, I'}  to denote that infrastructures 
\texttt{I} and \texttt{I'} are in this relation.
%Rules for the put, get, move, process and delete actions exist.
To give an impression of this definition, we show first the rule defining the state transition for
the action get. Initially, this rule expresses that
an actor that resides at a location \texttt{l} (\texttt{a \ttatI\ l})
and is enabled by the local policy in this location to ``get'' can
combine  all ids at the current location (contained in \texttt{egra G l}) with
all actors at the current location (contained in \texttt{agra G l}) and add this
set of pairs to his knowledge set \texttt{kgra G} using the function update \texttt{f(l := n)}
redefining the function \texttt{f} for the input \texttt{l} to have now the new value \texttt{n}.
\begin{ttbox}
 {\bf{get\_data}}: G = graphI I \ttImp a \ttatI l \ttImp enables I l (Actor a) get \ttImp 
   I' = Infrastructure 
          (Lgraph (gra G)(agra G)(cgra G)(lgra G)(egra G)
            ((kgra G)((Actor a) := ((kgra G (Actor a))(l:=
                 \{(x,y). x \ttin agra G l \ttand y \ttin egra G l\})))))
          (delta I)
   \ttImp I \ttrel{n} I' 
\end{ttbox}
Another interesting rule for the state transition is the one for \texttt{move} whose structure
resembles the previous one.
\begin{ttbox}
 {\bf{move}}: G = graphI I \ttImp a \ttatI l \ttImp a \ttin actors_graph(graphI I) \ttImp
    l \ttin nodes G \ttImp l' \ttin nodes G \ttImp enables I l' (Actor h) move \ttImp
    I' = Infrastructure (move_graph_a a l l' (graphI I))(delta I)
 \ttImp I \ttrelI I' 
\end{ttbox}
The semantics of this rule is embedded in the function \texttt{move\_graph\_a}
that adapts the infrastructure state so that the moving actor \texttt{a} is now
associated to the target location \texttt{l'} in the actor map \texttt{agra} and not
any more at \texttt{l} and also the association of \texttt{efids} is updated accordingly.
\begin{ttbox}
move_graph_a n l l' g \ttequiv
  Lgraph (gra g) 
         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then 
           ((agra g)(l := (agra g l) - \{n\}))(l' := (insert n (agra g l')))
          else (agra g))
         (cgra g)(lgra g)
         (if n \ttin ((agra g) l) \ttand  n \ttnin ((agra g) l') then
               ((egra g)(l := (egra g l) - \{efemid (cgra g n)\}))
                        (l' := (insert (efemid (cgra g n))(egra g l')))
          else egra g)
         (kgra g)
\end{ttbox}  
Based on this state transition and the above defined \texttt{Corona\_scenario}, we define the first Kripke structure.
\begin{ttbox}
 corona_Kripke \ttequiv Kripke \{ I. Corona_scenario \ttrelIstar I \} \{Corona_scenario\}
\end{ttbox}

@@

Since we consider a predicate transformer semantics, we use
sets of states to represent properties. 
%For example, the attack property
The invalidated global policy is given by the set \texttt{scorona}.
\begin{ttbox}
 scorona \ttequiv \{x. \ttexists n. \ttneg global_policy x (Efid n)\}
\end{ttbox}
The attack we are interested in is to see whether for the scenario
\begin{ttbox}
 Kripke\_scenario \ttequiv  Infrastructure ex_graph local_policies 
\end{ttbox}
from the initial state \texttt{Icorona \ttequiv \{corona\_scenario\}},
the critical state \texttt{scorona} can be reached,
that is, is there a valid attack \texttt{(Icorona,scorona)}?

For the Kripke structure \texttt{corona\_Kripke}
we first derive a valid and-attack using the attack tree proof calculus.
\begin{ttbox}
  \ttvdash [\ttcalN{(Icorona,Corona)},\ttcalN{(Corona,Corona1)},\ttcalN{(Corona1,Corona2)},
             \ttcalN{(Corona2,Corona3)},\ttcalN{(Corona3,scorona)}]\ttattand{\texttt{(Icorona,scorona)}}
\end{ttbox}
The sets \texttt{Corona, Corona1, Corona2, Corona3} are the intermediate states where
\texttt{Bob} moves to shop and \texttt{Eve} follows him collecting the Ephemeral Ids in each location.
The collected information enables identifying Bob's Ephemeral Id.

@@

\subsection{Refining the Specification}
\label{sec:corref}
Clearly, fixed Ephemeral Ids are not really ephemeral. The model presented
in Section \ref{sec:model} has deliberately been designed abstractly to allow focusing on
the basic system architecture and finding an initial deanonymization attack.
We now introduce ``proper'' Ephemeral Ids and show how the system datatype can be refined to a system
that uses those instead of the fixed ones.

@@

The first step of refinement replaces the simple \texttt{efid} in the infrastructure type by the
new type \texttt{efidlist}. Note, that in the new datatype \texttt{igraph} this change affects only
the thrird component, the credentials. The last two componenets, the set of currently present efids,
and the knowledge set, still operate on the simple type \texttt{efid}.
\begin{ttbox}
{\bf datatype} igraph = 
         Lgraph (location \tttimes location)set 
                 location \ttfun identity set
                 identity \ttfun (string set \tttimes string set \tttimes efidlist)  
                 location \ttfun string \tttimes (dlm \tttimes data) set
                 location \ttfun efid set
                 actor \ttfun location \ttfun (identity \tttimes efid) set
\end{ttbox}
The refined state transition relation implements the possibility of changing the Ephemeral Ids
by the rule for the action \texttt{put}.
\begin{ttbox}
 {\bf{put}}: G = graphI I \ttImp  a \ttatI l \ttImp enables I l (Actor a) put \ttImp
      I' = Infrastructure (put_graph_efid A l (graphI I))(delta I)
 \ttImp I \ttrelI I' 
\end{ttbox}
The change to the infrastructure state is implemented in the function \texttt{put\_graph\_efid} that
increases the current index in the \texttt{efidlist} in the credential component \texttt{cgra g n} for
the ``putting'' actor identity \texttt{n} and inserts the current \texttt{efid} from that credential
component into the \texttt{egra} component, the set of currently ``present'' Ephemeral Ids at the location
\texttt{l}.
\begin{ttbox}
put_graph_efid n l g  \ttequiv
  Lgraph (gra g)(agra g)
         ((cgra g)(n := (credentials (cgra g n), roles (cgra g n),
                         efids_inc_ind(efemid (cgra g n)))))
         (lgra g)
         ((egra g)(l := insert (efids_cur(efemid (cgra g n)))(egra g l)))
         (kgra g)
\end{ttbox}
We can now apply the refinement by defining a datatype map from the refined infrastructure type
\texttt{InfrastructureOne.infrastructure} to the former one \texttt{Infrastructure.infrastructure}.
\begin{ttbox}
{\bf{definition}} refmap :: InfrastructureOne.infrastructure \ttfun
                          Infrastructure.infrastructure
{\bf{where}} refmap I =
  Infrastructure.Infrastructure 
    (Infrastructure.Lgraph
       (InfrastructureOne.gra (graphI I))
       (InfrastructureOne.agra (graphI I))
        (\ttlam h. repl_efr 
           (InfrastructureOne.cgra (graphI I)) h)
           (InfrastructureOne.lgra (graphI I))
           (InfrastructureOne.egra (graphI I))
           (\ttlam a. \ttlam l.
  (\ttlam (x,y).(x, efids_root(efemid(InfrastructureOne.cgra (graphI I) x))))
                       \ttimg(InfrastructureOne.kgra (graphI I)) a l))                                                    
\end{ttbox}  
This is then plugged into the parameter \texttt{\ttecal} of the refinement operator allowing to
prove \texttt{corona\_Kripke \ttref{\texttt{refmap}} corona\_KripkeO} where the latter is the
refined Kripke structure.

Surprisingly, we can still prove \texttt{corona\_KripkeO \ttvdash EF scoronaO} by using the
same attack tree as in the abstract model: if Bob moves from pub to shop, he is vulnerable to
being identifiable as long as he does not change the current \texttt{efid}. So, if Eve moves to
the shop as well and performs a get before Bob does a put, then Eve's knowledge sets permits identifying
Bob's current Ephemeral Id as his.

\subsubsection{Second refinement step}
The persistent attack can be abbreviated informally by the action sequence \texttt{[get,move,move,get]}
performed by actors Eve, Bob, Eve, and Eve again, respectively.
How can a second refinement step avoid that Eve does the last \texttt{get} by imposing that after the
first \texttt{move} of Bob a \texttt{put} action must happen before Eve can do another \texttt{get}?
A very simple remedy to exclude this attack is to bind a \texttt{put} action after every \texttt{move}.
We can implement that change by a minimal update to the function \texttt{move\_graph\_a} (see Section \ref{sec:model})
by adding an increment (highlighted in the code snippet)
of the currently used Ephemeral Id before updating the \texttt{egre} component of the target location.
\begin{ttbox}
move\_graph\_a n l l' g \ttequiv ...
  (l' := insert (efids_cur({\bf efids\_inc\_ind}(efemid (cgra g n))))(egra g l))...
\end{ttbox}
